---
title: "Google Analytics Sample - Data Extraction"
author: "Tung Anh"
date: "2025-07-16"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Introduction
This notebook shows how to extract and process \href{https://console.cloud.google.com/marketplace/product/obfuscated-ga360-data/obfuscated-ga360-data?q=search&referrer=search&inv=1&invt=Ab26RA&project=bigdata-382912}{\textbf{Google Analytics sample}} data from BigQuery API. The dataset contains 12 months of anonymized data (August 2016 to August 2017) from the Google Merchandise Store, a real ecommerce site selling Google-branded merchandise.

# Setup

```{r pkg.install, include=FALSE}

# install.packages("bigrquery")      
# install.packages("DBI")            
# install.packages("plotly")         
# install.packages("dplyr")          
# install.packages("tidyr")          
# install.packages("readr")          
# install.packages("bigrquery")
# install.packages("lubridate")
# install.packages("arrow")
# install.packages("glue")
```

```{r pkg.load}
# 📦 Load required libraries
library(bigrquery)
library(DBI)
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(arrow)
library(glue)
```

# Authentication

To use Big Query API directly from a Google Account, we use web browser authentication:
```{r}
bq_auth() # Sign in to Google Account with active Google Cloud console
```

# Original data

The original data can be access using the code below:

```{r}
# Data importation tester
project_id <- "bigdata-382912"
sql <- "
SELECT * 
FROM `bigquery-public-data.google_analytics_sample.ga_sessions_20170801` 
LIMIT 100"

tb <- bq_project_query(project_id, sql)
org_df <- bq_table_download(tb)

head(org_df)
```
The original data contains 16 columns, 6 of them are nested (`totals`, `trafficSource`, `device`, `geoNetwork`, `customDimensions`, `hits`) and 3 of them are obfuscated (`visitorId`, `userId` and `clientId`).

# Extract one day of flattened columns
We access 1 day of data from Google Analytic Sample dataset, by flatten a selection of unnested columns and also neglected obfuscated columns:

```{r}
# Set your project ID
project_id <- "bigdata-382912"

# Write your SQL query (flattened version to avoid nested errors)
main_query <- "
SELECT
  -- single columns
  date,
  fullVisitorId,
  visitId,
  visitNumber,
  visitStartTime,
  channelGrouping,
  -- total unnested columns
  totals.newVisits AS total_newVisits,
  totals.pageviews AS total_pageviews,
  totals.hits AS total_hits,
  totals.bounces AS total_bounces,
  totals.timeOnsite AS total_timeOnsite,
  totals.transactions AS total_transactions,
  totals.transactionRevenue AS total_transactionRevenue,
  totals.totalTransactionRevenue AS total_totalTransactionRevenue,
  -- trafficSource unnested columns
  trafficSource.source AS trafficSource_source,
  trafficSource.campaign AS trafficSource_campaign,
  -- device unnested columns
  device.browser AS device_browser,
  device.deviceCategory As device_deviceCategory,
  -- geoNetwork unnested columns
  geoNetwork.continent as geoNetwork_continent,
  geoNetwork.country AS geoNetwork_country,
  
FROM `bigquery-public-data.google_analytics_sample.ga_sessions_20170101`
"

# Run the query and save to a dataframe
df <- bq_project_query(project_id, main_query) %>%
      bq_table_download()

# View the result
head(df)
```

**Data description**

- `date`: Date of the session in YYYYMMDD format.  
- `fullVisitorId`: Unique identifier for a user (visitor) across multiple sessions.  
- `visitId`: Unique identifier for a specific session (visit) within a user’s visits.  
- `visitNumber`: Number of visits a user has made to the site up to and including this session.  
- `visitStartTime`: Timestamp (in UNIX format) when the session started.  
- `channelGrouping`: Traffic channel category that brought the visitor to the site (e.g., Organic Search, Direct, Referral).  
- `total_newVisits`: Indicator if this session was the user's first visit (1 if new visitor, else 0).  
- `total_pageviews`: Total number of pages viewed during the session.  
- `total_hits`: Total hits (interactions) recorded during the session.  
- `total_bounces`: Number of single-page sessions (bounces) during this visit.  
- `total_timeOnsite`: Total time spent on the site during the session (in seconds).  
- `total_transactions`: Number of transactions completed during the session.  
- `total_transactionRevenue`: Total revenue generated from transactions during the session (in micros, i.e., multiplied by 1,000,000).  
- `total_totalTransactionRevenue`: Total transaction revenue aggregated across all transactions during the session (in micros, i.e., multiplied by 1,000,000). 
- `trafficSource_source`: Source of the traffic for the session (e.g., google, direct, newsletter).  
- `trafficSource_campaign`: Marketing campaign name associated with the traffic.  
- `device_browser`: Browser used by the visitor (e.g., Chrome, Firefox).  
- `device_deviceCategory`: Device category used to access the site (e.g., desktop, mobile, tablet).  
- `geoNetwork_continent`: Continent from which the session originated.  
- `geoNetwork_country`: Country from which the session originated.  

# Generalized function to extract data by range of date

This function downloads daily data for a given date range and saves it as Parquet files organized by year and month:

```{r data_download_function}
download_ga_data <- function(project_id, start_date, end_date, saved_path, cardinality_threshold = 10) {
  # Parse start and end dates
  start_date <- ymd(start_date)
  end_date <- ymd(end_date)
  
  current_date <- start_date
  
  while (current_date <= end_date) {
    date_str <- format(current_date, "%Y%m%d")
    year_str <- format(current_date, "%Y")
    month_str <- format(current_date, "%m")
    
    # Build query with glue
    main_query <- glue("
      SELECT
        date,
        fullVisitorId,
        visitId,
        visitNumber,
        visitStartTime,
        channelGrouping,
        totals.newVisits AS total_newVisits,
        totals.pageviews AS total_pageviews,
        totals.hits AS total_hits,
        totals.bounces AS total_bounces,
        totals.timeOnsite AS total_timeOnsite,
        totals.transactions AS total_transactions,
        totals.transactionRevenue AS total_transactionRevenue,
        totals.totalTransactionRevenue AS total_totalTransactionRevenue,
        trafficSource.source AS trafficSource_source,
        trafficSource.campaign AS trafficSource_campaign,
        device.browser AS device_browser,
        device.deviceCategory AS device_deviceCategory,
        geoNetwork.continent AS geoNetwork_continent,
        geoNetwork.country AS geoNetwork_country
      FROM `bigquery-public-data.google_analytics_sample.ga_sessions_{date_str}`
    ")
    
    # Query and download
    df <- bq_project_query(project_id, main_query) %>% bq_table_download()
    
    # Convert numeric columns with low cardinality to factor
    df <- df %>%
      mutate(across(
        .cols = where(is.numeric),
        .fns = ~ if (n_distinct(.) <= cardinality_threshold) factor(.) else .
      ))
    
    # Create directories if needed
    dir_path <- file.path(saved_path, year_str, month_str)
    if (!dir.exists(dir_path)) dir.create(dir_path, recursive = TRUE)
    
    # Save as parquet
    save_path <- file.path(dir_path, paste0("ga_sessions_", date_str, ".parquet"))
    arrow::write_parquet(df, save_path)
    
    message(glue("Saved data for {date_str} to {save_path}"))
    
    # Next day
    current_date <- current_date + days(1)
  }
}
```

# Full download

To download full data (from August 2016 to August 2017), we run the code below:

```{r full_download}
download_ga_data(
  project_id="bigdata-382912",
  start_date="20170101",
  end_date="20170102", 
  saved_path = "/data/")
```




























